\vspace{-1cm} Model simplicity breeds contempt: using simple models to answer basic questions on species' distributions

### 2021-02-10 manuscript/manuscript.tex
Trying again with logistic... this sucks

One of the central goals of ecology is to understand the ways species are distributed across space (ref). Over the years, researchers have developed multiple distribution models to try to untangle the factors that play a role in defining such distributions (ref). These models estimate species' realized niches using several covariates, including environmental variables (ref), species ecological traits' (ref) and phylogenetic relations (ref). Recent work on these approaches has focused on estimating (ref) and accounting for (ref) biotic factors, such as competitive or facilitative relationships. The idea is that by understanding how such biotic and abiotic factors shape species' distributions, we can gain a mechanistic understanding on how ecological communities are established and change over time. Unfortunately, while some of this approaches can increase the predictive performance of distribution models (ref), the nature of some of the estimates have been theoretically questioned (ref). For example...

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PARAGRAPH 3
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
While the value of gaining a mechanistic understanding of species' distributions is unquestioned, a lot can be learned from shedding light on basic properties of species' realized niches.

@@

@@

To decide among modelling approaches, we first needs to agree on what we know about the system. We know that species occupy a geographic range; therefore, we know that their distributions have finite variance. Indeed, observations on species' geographic variation and optimal climatic conditions have been long documented, with extensive databases compiled by botanists and field ecologists documenting basic knowledge on species' distributions. One could point out that we also know that many other factors might influence species' presence/absence---e.g. the influence of biotic interactions among species. However, we do not necessarily have an intuition of how exactly these factors will influence the shape of species' distributions. As a result, if all we truly knew about a species' distribution was that they have finite variance, the most conservative assumption and the safest bet---i.e. the one with the largest entropy---is that such distribution is a Gaussian.

>>>>

One of the central goals of ecology is to understand the ways species are distributed across space and time (ref). As a result, ecologists have developed multiple distribution models over the last two decades, to try to untangle the factors that play a role in defining such distributions (ref). These models estimate species' realized niches using several covariates, including environmental variables (ref), species ecological traits' (ref) and phylogenetic relations (ref). For example, ... More recently, some of the focus have shifted towards approaches that estimate and account for biotic factors, such as competitive or facilitative relationships between species. The idea is that by understanding how such biotic and abiotic factors shape species' distributions, we can gain a mechanistic understanding on how ecological communities are established and change over time. However, while some of this approaches can increase the predictive performance of distribution models (ref), the nature of some of the estimates have been theoretically questioned (ref). This was parsimoniously illustrated by \citealt{bla bla}, who ...

%Comparisons of the predictive performance of such models are also extensive (ref), untangling the modelling approaches that better capture variation in species' distributions.

% People comparing performance: Wilkinson, D. P., N. Golding, G. Guillera‐Arroita, R. Tingley, and M. A. McCarthy. 2019. A comparison of joint species distribution models for presence–absence data. Methods in Ecology and Evolution 10:198–211; A comprehensive evaluation of predictive performance of 33 species distribution models at species and community levels

%The last two decades have seen a proliferation of species distribution models (SDMs) addressing the challenge of predicting the occurrences of individual species (Guisan and Zimmermann 2000, Guisan and Thuiller 2005, Elith et al. 2006, Leathwick et al. 2006, Zimmermann et al. 2010). Methodological advances in multiple‐species distribution modeling have lagged behind, but are recently experiencing a rapid expansion (Leathwick et al. 2006, Dunstan et al. 2011, Guisan and Rahbek 2011, Warton et al. 2015, Wilkinson et al. 2019). Many previous studies (Table 1) have compared the predictive performance of SDMs for single‐species analyses (Moisen and Frescino 2002, Thuiller et al. 2003, Elith et al. 2006, Leathwick et al. 2006, Elith and Graham 2009, Guisan and Rahbek 2011). Some studies have compared single‐species and multi‐species distribution models (Araújo and Luoto 2007, Heikkinen et al. 2007, Baselga and Araújo 2009, 2010, Elith and Leathwick 2009, Chapman and Purse 2011, Bonthoux et al. 2013, Madon et al. 2013, Maguire et al. 2016, Harris et al. 2018), while a few have examined the performance of alternative multiple species modeling approaches (Baselga and Araújo 2010, Madon et al. 2013, Wilkinson et al. 2019).

Unfortunately, while some of this approaches can increase the predictive performance of distribution models (ref), the nature of some of the estimates have been theoretically questioned (ref). For example...

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PARAGRAPH 3
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% From Ana Norberg: Each statistical modeling method is based on different assumptions that can be viewed as hypotheses about how ecological communities are structured

While the value of gaining a mechanistic understanding of species' distributions is unquestioned, a lot can be learned from shedding light on basic properties of species' realized niches.

@@

 For non-parametric models, these often do not account for species uncertainties. One can benefit a lot from a bayesian framework. First, it allows us to compare the parameter estimates ... It also help us incorporate prior information.

@@

To decide among modelling approaches, we first need to agree on what we know about the system. We know that species occupy a geographic range; therefore, we know that their distributions have finite variance. Indeed, observations on species' geographic variation and optimal climatic conditions have been long documented, with extensive databases compiled by botanists and field ecologists documenting basic knowledge on species' distributions. One could point out that we also know that many other factors might influence species' presence/absence---e.g. the influence of biotic interactions among species. However, we do not necessarily have an intuition of how exactly these factors will influence the shape of species' distributions. As a result, if all we truly knew about a species' distribution was that they have finite variance, the most conservative assumption and the safest bet---i.e. the one with the largest entropy---is that such distribution is a Gaussian.
